---
title: "Avocados Pricing Engine: Capstone Project"
subtitle: "Data Science Professional Certificate Program (HarvardX - EdX Platform)"
author: "Luis Sousa"
date: "December 2021"
output:
  html_document:
    dev: png
    toc: yes
    fig_width: 10 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. INTRODUCTION

l
https://www.kaggle.com/janiobachmann/price-of-avocados-pattern-recognition-analysis


ABOUT THE DATA: https://www.kaggle.com/neuromusic/avocado-prices

## 1.1 Goals & Challenges

l

* l
* l

l

# 2. DATA ANALYSIS

l

```{r echo=F, results='hide', eval=T, error=F, warning=F, message=F}
##########################################################
# Create edx set, validation set (final hold-out test set)
##########################################################

# Note: this process could take a couple of minutes
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org", dependencies = T)
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org", dependencies = T)
if(!require(data.table)) install.packages("data.table", repos = "http://cran.us.r-project.org", dependencies = T)
if(!require(stringi)) install.packages("stringi", repos = "http://cran.us.r-project.org", dependencies = T)
if(!require(dplyr)) install.packages("dplyr", repos = "http://cran.us.r-project.org", dependencies = T)
if(!require(knitr)) install.packages("knitr", repos = "http://cran.us.r-project.org", dependencies = T)
if(!require(kableExtra)) install.packages("kableExtra", repos = "http://cran.us.r-project.org", dependencies = T)
if(!require(ggplot2)) install.packages("ggplot2", repos = "http://cran.us.r-project.org", dependencies = T)
if(!require(gridExtra)) install.packages("gridExtra", repos = "http://cran.us.r-project.org", dependencies = T)
if(!require(lubridate)) install.packages("lubridate", repos = "http://cran.us.r-project.org", dependencies = T)
if(!require(corrgram)) install.packages("corrgram", repos = "http://cran.us.r-project.org", dependencies = T)

library(tidyverse)
library(caret)
library(data.table)
library(stringi)
library(dplyr)
library(knitr)
library(ggplot2)
library(gridExtra)
library(lubridate)
library(corrgram)
```

## 2.1 Downloading Data

ABC

```{r echo=T, results='hide', eval=T, error=F, warning=F, message=F}
# Original Kaggle link with the Avocado dataset
Kaggle_data_link <- 
    "https://www.kaggle.com/neuromusic/avocado-prices/download/"

# Mirrow Avocado dataset on gitHub
gitHub_data_link <- 
  "https://github.com/luisfdsousa/EdX_Capstone_Avocado/raw/main/avocado.csv"

# Download link
data_weblink <- gitHub_data_link

# File name containing the desired dataset
data_file_name <- "avocado.csv"

# Create an empty avocados data frame
avocados <- NULL

#Check if data exists within the working directory
if (file.exists(data_file_name) == F) {
  
  #Info message
  cat("Avocado data not found.",
      "Next step: Download data from Kaggle")
  
  # Create temporary file to store the data
  temp_file <- tempfile()
  
  # Download the data from the web
  download.file(data_weblink, temp_file)
  
  # Open and load .cvs data 
  avocados <- read.csv(temp_file,
                       header = T)
  
  # Free memory, remove temp file
  rm(temp_file)
  
} else {
  
  # Open and load .cvs data 
  avocados <- read.csv(data_file_name,
                       header = T)
}
```

## 2.2 Data Transformation & Cleaning

```{r echo=T, results='hide', eval=T, error=F, warning=F, message=F}
# Drop the first column (gibberish)
# drop also the bags column (description not clear)
avocados <- 
  avocados %>%            # pipe the avocados data set
  subset(select = -c(X,
                     Total.Bags,
                     Small.Bags,
                     Large.Bags,
                     XLarge.Bags))

# Rename columns to simplify reading
names(avocados)[names(avocados) == "Total.Volume"] <- "TotalVolume"
names(avocados)[names(avocados) == "X4046"] <- "T4046"
names(avocados)[names(avocados) == "X4225"] <- "T4225"
names(avocados)[names(avocados) == "X4770"] <- "T4770"
names(avocados)[names(avocados) == "type"] <- "Type"
names(avocados)[names(avocados) == "year"] <- "Year"
names(avocados)[names(avocados) == "region"] <- "Region"

#avocados$Year = as.factor(avocados$Year)
avocados$Type = as.factor(avocados$Type)
avocados$Region = as.factor(avocados$Region)
avocados$Year = as.factor(avocados$Year)
#avocados$Date = as.Date(avocados$Date)
avocados$Date = ymd(avocados$Date)
avocados <- add_column(avocados, Month = factor(months(avocados$Date), levels = month.name))
avocados <- add_column(avocados, Day = factor(format(avocados$Date, format = "%d")))
avocados <- avocados %>% mutate(NonHaasVolume = TotalVolume - T4046 - T4225 - T4770)

# Remove NA entries
avocados <- avocados %>% drop_na()
```

Data Cleaning

```{r echo=T, results='hide', eval=T, error=F, warning=F, message=F}
# Initiate seed for repeated results
# if using R 3.5 or earlier, use `set.seed(1)`
set.seed(1, sample.kind = "Rounding")

# 90/10 % ratio for test/validation datasets
training_samples <- 
  avocados$AveragePrice %>%
  createDataPartition(p = 0.9, 
                      list = FALSE)

# Create respective train and test datasets 
train_data  <- avocados[training_samples, ]
test_data <- avocados[-training_samples, ]
```

## 2.3 Basic Data Exploration: R Basic Functions

```{r echo=T, results=T, eval=T, error=F, warning=F, message=F}
# Summary of the data
data_basics_metrics <- c(
  nrow(avocados),
  ncol(avocados),
  anyNA(avocados),
  n_distinct(avocados$Type),
  n_distinct(avocados$Region),
  min(year(avocados$Date)),
  max(year(avocados$Date)))

data_basics_description <- c(
  "Number of Rows",
  "Number of Columns",
  "Number of Invalid Data Points",
  "Distinct Types of Avocados",
  "Distinct Produce Regions",
  "Data Start Date",
  "Data End Date"
)

# Build data frame with columns + description
data_basics <- data.frame(data_basics_description,
                          data_basics_metrics)

# Name the columns of the table
names(data_basics)<-c("Metric","Result")

# Draw the table
data_basics %>%
  kbl(caption = "Basic Data Exploration: Avocado Dataset") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

Comments

```{r echo=T, results=T, eval=T, error=F, warning=F, message=F}
# Save column names
avocados_col <- colnames(avocados)

# Add column meaning
avocados_col_meaning <- c(
  "The date of the observation",
  "The average price of a single avocado",
  "Total number of avocados sold",
  "Total number of avocados: Small/medium Haas",
  "Total number of avocados: Large Haas",
  "Total number of avocados: Extra large Haas",
  "Type of avocado",
  "Produce year",
  "Produce region"
)

# Build data frame with columns + description
results <- data.frame(avocados_col[1:(length(avocados_col) - 3)], 
                      avocados_col_meaning)

# Name the columns of the table
names(results) <- c("Columns", "Description")

# Draw the table
results %>%
  kbl(caption = "Avocado dataset: Columns description") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```


```{r echo=T, results=T, eval=T, error=F, warning=F, message=F}
# Print data summary
summary(avocados)
```

## 2.4 Advanced Data Exploration: Data Visualization

bla

### 2.4.1 Plot 1: Histograms of Average Avocado Prices

```{r echo=F, results=T, eval=T, error=F, warning=F, message=F}
# Plot #1: Histograms of Average Avocado Prices
par(mar=c(1,1,1,1))

# Calculate mean average  prices per type of avocado
mean_avg_prices <- avocados %>%
  group_by(Type) %>%
  summarise(avg_price = mean(AveragePrice))

# Average Prices histogram
plot1_hist1 <- avocados %>% 
  ggplot(aes(x = AveragePrice, color = Type)) + 
  geom_histogram(binwidth = 0.05, position = "dodge", fill = "white", alpha = 0.5) + 
  facet_grid(Type ~ .) +
  geom_vline(data = mean_avg_prices, aes(xintercept = avg_price, color = Type), linetype = "dashed") + 
  labs(title = "Average Avocado Price", x = "Average Price per Avocado", y = "Count") + 
  theme_classic() + 
  theme(legend.position = "bottom")

# Change line colors by groups
plot1_hist2 <- avocados %>% 
  ggplot(aes(x = AveragePrice, color = Type, fill = Type)) +
  geom_histogram(aes(y = ..density..), binwidth = 0.05, position = "dodge", alpha = 0.2) +
  facet_grid(Type ~ .) +
  geom_density(alpha = 0.3) +
  geom_vline(data = mean_avg_prices, aes(xintercept = avg_price, color = Type), linetype = "dashed") +
  labs(title = "Average Avocado Price (Density)", x = "Average Price per Avocado", y = "Density") +
  theme_classic() + 
  theme(legend.position = "bottom")

# Display plots
grid.arrange(plot1_hist1, plot1_hist2, ncol = 2) # Apply grid.arrange function
```

bla bla

### 2.4.2 Plot 2: Histograms of Average Avocado Prices

bla bla

```{r echo=F, results=T, eval=T, error=F, warning=F, message=F}
# Plot #2: Histograms of Average Avocado Prices
plot2_boxplot_price <- avocados %>%
  ggplot(aes(Type, AveragePrice, color = Year)) +
  geom_boxplot() +
  labs(title = "Average Avocado Price", x = "Type of Avocado", y = "Average Price per Avocado") +
  theme_classic() + 
  theme(legend.position = "bottom")

plot2_boxplot_volume <- avocados %>% 
  ggplot(aes(Type, TotalVolume, color = Year)) +
  geom_boxplot() +
  scale_y_continuous(trans='log2') +
  labs(title = "Total Volume Sold", x = "Type of Avocado", y = "Total Volume Sold") +
  theme_classic() + 
  theme(legend.position = "bottom")

# Display plots
grid.arrange(plot2_boxplot_price, plot2_boxplot_volume, ncol = 2) # Apply grid.arrange function
```

bla bla

### 2.4.3 Plot 3: Total volume sold by Type & Year

bla bla

```{r fig.height = 8, echo=F, results=T, eval=T, error=F, warning=F, message=F}
# Plot #3 Total volume sold by Type & Year
plot3_scatterplot_volPrice <- avocados %>%
  ggplot(aes(TotalVolume, AveragePrice, colour = Type)) +
  geom_point(alpha = 0.05) + 
  geom_smooth(color = "black", method = "lm", linetype = "dashed") +
  scale_x_continuous(trans = 'log2') +
  facet_grid(Type ~ .) +
  labs(title = "Average Avocado Price", x = "Total Volume Sold", y = "Average Price per Avocado") +
  theme_classic()

plot3_lineplot_TotalMonthlyRevenue <- avocados %>%
  group_by(Year, Type, Month) %>%
  select(AveragePrice, Year, Month, Type, TotalVolume) %>%
  mutate(MonthlyRevenue = AveragePrice*TotalVolume) %>%
  summarise(TotalMonthlyRevenue = sum(MonthlyRevenue)) %>%
  ggplot(aes(x = Month, y = TotalMonthlyRevenue, colour = Year, group = Year)) +
  geom_line() + 
  facet_grid(. ~Type) + 
  scale_y_continuous(trans = 'log2') +
  labs(x = "Month", y = "Monthly Revenue (in mn)", title = "Average Monthly Avocado Prices") +
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 65, hjust = 1))

# Display plots
grid.arrange(plot3_scatterplot_volPrice, plot3_lineplot_TotalMonthlyRevenue, nrow = 2)
```

bla bla

### 2.4.4 Plot 4: Total Rev sold by Size, Year & Type

bla bla

```{r echo=F, results=T, eval=T, error=F, warning=F, message=F}
# Plot #4: Total Rev sold by Size, Year & Type
plot4 <- avocados %>%
  gather(AvocadoSize, Volume, c(T4046:T4770,NonHaasVolume)) %>%
  mutate(Revenue = AveragePrice * Volume) %>% 
  group_by(Year, Month, Type, AvocadoSize) %>%
  summarise(TotalRevenue = sum(Revenue)/1000000) %>%
  ggplot(aes(x = Month, y = TotalRevenue, colour = Year, group = Year)) +
  facet_grid(vars(Type),vars(AvocadoSize), scales="free") +
  geom_line() +
  scale_y_continuous(trans = 'log2') +
  labs(y = "Monthly Revenue (in mn)", x ="Month", title = "Average Monthly Avocado Prices by Size", colour = 'Year') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 55, hjust = 1),
        panel.grid.minor = element_line(size = 0.4, linetype = 1),
        panel.grid.major = element_line(size = 0.75, linetype = 3))

# Display plot
plot4
```

bla bla


### 2.4.5 Plot 5: Average Price per date: Conventional/organic Avocados

bla bla

```{r echo=F, results=T, eval=T, error=F, warning=F, message=F}
# Plot #5: Average Price per date: Conventional/organic Avocados
plot5 <- avocados %>%
  ggplot(aes(x = Date, y = AveragePrice, color = Type)) +
  geom_line(alpha = 0.8) +
  theme_classic() +
  scale_x_date(date_breaks = "6 month", date_labels =  "%b %Y", expand = c(0,0)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

# Display plot
plot5
```

bla

### 2.4.6 Plot 6: Total Rev sold by Size, Year & Type

bla bla

```{r fig.height = 10, echo=F, results=T, eval=T, error=F, warning=F, message=F}
# Plot #6: Total Rev sold by Size, Year & Type
plot6 <-  avocados %>% 
  ggplot(aes(x = Region, y = AveragePrice, colour = Type)) + #, group = Year
  geom_boxplot() + 
  facet_grid(.~Year, scales = "free") +
  scale_y_continuous(breaks = c(seq(round(min(avocados$AveragePrice),1),
                                    round(max(avocados$AveragePrice),1),0.2)), 
                     limits = c(round(min(avocados$AveragePrice),1),
                                round(max(avocados$AveragePrice),1))) +
  labs(x = "Produce Region", y ="Average Avocado Price", 
       title = "Average prices by region & year") +
  coord_flip() + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0),
        panel.grid.minor = element_line(size = 0.4, linetype = 1),
        panel.grid.major = element_line(size = 0.75, linetype = 3))

# Display plot
plot6
```

bla

### 2.5.4 Remarks

When it comes to to the analysis we have implemented there are several observations we have observed:

* Expensive Organic Avocados: As expected, we have noticed that organic avocados are much more expensive than conventional avocados.
* Similar patterns by type: There are some distinguishable patterns between the two types of avocados however, we see that most patterns are similar between these two types of avocados.
* Year 2017: The year of 2017 was the best year of avocados. One of the factors that could have contributed was the well being of the economy as a whole but there could also be other hidden factors that could affect avocado prices in the market.
* Volatility: Although, 2017 was the year with the highest prices, it was also the year with the highest volatility. This means that price fluctuations were higher in this year as oppose to the past two years.
* Buy avocados before fall!: We see a consistent pattern of avocado prices increasing when fall is coming. This applies to both conventional and organic avocados.
* Downward trend in the longrun: Based on our ARIMA model, for both types of avocados we expect a downward trend in prices at least in the longrun. There could be some upward pressure in the short run however, our model predicts that both avocado prices will o downwards 2 years from now.

Source: https://www.kaggle.com/janiobachmann/price-of-avocados-pattern-recognition-analysis

# 3. METHODS: MODEL ENGINEERING

l

## 3.1 Model 1: Average Model

BLA

```{r echo=T, results='hide', eval=T, error=F, warning=F, message=F}
RMSE <- function(observed_ratings, expected_ratings){
  sqrt(mean((observed_ratings - expected_ratings)^2))
}
```

BLA

```{r echo=T, results=T, eval=T, error=F, warning=F, message=F}
# Compute the dataset's mean rating
mean_avocado_price <- mean(train_data$AveragePrice)
cat("The overall mean avocado price is: ", round(mean_avocado_price, 2))

# Validate model 1
model1_mean <- RMSE(test_data$AveragePrice, mean_avocado_price)
cat("RMSE =", model1_mean, "Assuming the same mean price", round(mean_avocado_price, 2), " for all avocados")
```

BLA

```{r echo=F, results=T, eval=T, error=F, warning=F, message=F}
# MODEL 1: Print results
train_data %>%
  ggplot(aes(x = Date, y = AveragePrice)) + 
  geom_point(color = "gray", alpha = 0.1) +
  scale_x_date(date_breaks = "3 month", date_labels =  "%b %Y", expand = c(0,0)) +
  geom_hline(yintercept = mean_avocado_price, 
             linetype = 1, 
             size = 1) +
  labs(title = "MODEL 1: Average Model", x = "Date", y = "Average Avocado Price") + 
  theme_minimal() +
  geom_text(aes(median(Date), 
                round(mean_avocado_price, 2), 
                label = round(mean_avocado_price, 2),
                vjust = - 1))
```

BLA

## 3.2 Model 2: Linear Regression

BLA

```{r echo=F, results=T, eval=T, error=F, warning=F, message=F}
# Correlation matrix: Understand variables correlation
corrgram(avocados, order = TRUE, upper.panel = panel.cor, 
         main = "Correlation matrix: Avocados Dataset")
```

What Is R-squared?

R-squared is a statistical measure of how close the data are to the fitted regression line. It is also known as the coefficient of determination, or the coefficient of multiple determination for multiple regression.

The definition of R-squared is fairly straight-forward; it is the percentage of the response variable variation that is explained by a linear model. Or:

R-squared = Explained variation / Total variation

R-squared is always between 0 and 100%:
 
* 0% indicates that the model explains none of the variability of the response data around its mean.
* 100% indicates that the model explains all the variability of the response data around its mean.

In general, the higher the R-squared, the better the model fits your data. However, there are important conditions for this guideline that I’ll talk about both in this post and my next post.

```{r echo=T, results=T, eval=T, error=F, warning=F, message=F}
# Train Model
model2_linear = lm(AveragePrice ~ TotalVolume, data = avocados)
model2_linear_summary <- summary(model2_linear)
cat("R-squared =", model2_linear_summary$r.squared, "considering a linear model as a function of the total volume sold")
model2_linear_summary

# Validate model
model2_linear_validation <- model2_linear %>% predict(test_data)
model2_RMSE <- RMSE(test_data$AveragePrice, model2_linear_validation)
cat("RMSE =", model2_RMSE, "considering a linear model as a function of the total volume sold")
```

BLA

```{r echo=T, results=T, eval=T, error=F, warning=F, message=F}
# MODEL 2: Print results
test_data %>% 
  mutate(model2_price_prediction = model2_linear_validation) %>%
  ggplot() + 
  geom_point(aes(x = Date, y = AveragePrice, color = Type)) + 
  geom_line(aes(x = Date, y = model2_price_prediction)) +
  scale_x_date(date_breaks = "6 month", date_labels =  "%b %Y", expand = c(0,0)) +
  labs(title = "MODEL 2: Linear Regression", x = "Date", y = "Average Avocado Price") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, vjust = 0),
        panel.grid.minor = element_line(size = 0.4, linetype = 1),
        panel.grid.major = element_line(size = 0.75, linetype = 3),
        legend.position = "bottom")
```

BLA

## 3.3 Model 3: Multi Linear Regression

BLA

```{r echo=F, results=T, eval=T, error=F, warning=F, message=F}
# Train Model: version 1 (TotalVolume + Type)
model3_multiLinear.1 = lm(AveragePrice ~ TotalVolume + Type, data = avocados)
model3_multiLinear.1.summary <- summary(model3_multiLinear.1)
model3_multiLinear.1.summary

# the lm function has taken conventional type as being 0

cat("R-squared =", model3_multiLinear.1.summary$r.squared, "considering a multi linear model (total volume & type) as a function of the total volume sold")

# Validate model: version 1 (TotalVolume + Type)
model3_multiLinear.1.validation <- model3_multiLinear.1 %>% predict(test_data)
model3_multiLinear.1.RMSE <- RMSE(test_data$AveragePrice, model3_multiLinear.1.validation)
cat("RMSE =", model3_multiLinear.1.RMSE, "considering a multi linear model as a function of the total volume sold")

# MODEL 3: Print results: version 1 (TotalVolume + Type)
model3_multiLinear.1.plot1 <- test_data %>% 
  mutate(model3_price_prediction = model3_multiLinear.1.validation) %>%
  ggplot() + 
  geom_point(aes(x = Date, y = AveragePrice, color = Type)) + 
  geom_line(aes(x = Date, y = model3_price_prediction)) +
  scale_x_date(date_breaks = "6 month", date_labels =  "%b %Y", expand = c(0,0)) +
  labs(title = "MODEL 3: Multi Linear Regression (TotalVolume + Type)", x = "Date", y = "Average Avocado Price") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, vjust = 0),
        panel.grid.minor = element_line(size = 0.4, linetype = 1),
        panel.grid.major = element_line(size = 0.75, linetype = 3),
        legend.position = "bottom")

# Train Model: version 2 (TotalVolume + Type + Month + Year)
model3_multiLinear.2 = lm(AveragePrice ~ TotalVolume + Type + Month + Year, data = avocados)
model3_multiLinear.2.summary <- summary(model3_multiLinear.2)
model3_multiLinear.2.summary

cat("R-squared =", model3_multiLinear.2.summary$r.squared, "considering a multi linear model (Total volume + Type + Month + Year) as a function of the total volume sold")

# Validate model: version 2 (TotalVolume + Type + Month + Year)
model3_multiLinear.2.validation <- model3_multiLinear.2 %>% predict(test_data)
model3_multiLinear.2.RMSE <- RMSE(test_data$AveragePrice, model3_multiLinear.2.validation)
cat("RMSE =", model3_multiLinear.2.RMSE, "considering a multi linear model as a function of the total volume sold")

# MODEL 3: Print results: version 2 (TotalVolume + Type + Month + Year)
model3_multiLinear.2.plot2 <- test_data %>% 
  mutate(model3_price_prediction = model3_multiLinear.2.validation) %>%
  ggplot() + 
  geom_point(aes(x = Date, y = AveragePrice, color = Type)) + 
  geom_line(aes(x = Date, y = model3_price_prediction)) +
  scale_x_date(date_breaks = "6 month", date_labels =  "%b %Y", expand = c(0,0)) +
  labs(title = "MODEL 3: Multi Linear Regression (TotalVolume + Type + Month + Year)", x = "Date", y = "Average Avocado Price") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, vjust = 0),
        panel.grid.minor = element_line(size = 0.4, linetype = 1),
        panel.grid.major = element_line(size = 0.75, linetype = 3),
        legend.position = "bottom")

# Train Model: version 3 (TotalVolume + Type + Month + Year + Region)
model3_multiLinear.3 = lm(AveragePrice ~ TotalVolume + Type + Month + Year + Region, data = avocados)
model3_multiLinear.3.summary <- summary(model3_multiLinear.3)
model3_multiLinear.3.summary

cat("R-squared =", model3_multiLinear.3.summary$r.squared, "considering a multi linear model (Total volume + Type + Month + Year + Region) as a function of the total volume sold")

# Validate model: version 3 (TotalVolume + Type + Month + Year + Region)
model3_multiLinear.3.validation <- model3_multiLinear.3 %>% predict(test_data)
model3_multiLinear.3.RMSE <- RMSE(test_data$AveragePrice, model3_multiLinear.3.validation)
cat("RMSE =", model3_multiLinear.3.RMSE, "considering a multi linear model as a function of the total volume sold")

# MODEL 3: Print results: version 3 (TotalVolume + Type + Month + Year + Region)
model3_multiLinear.3.plot3 <- test_data %>% 
  mutate(model3_price_prediction = model3_multiLinear.3.validation) %>%
  ggplot() + 
  geom_point(aes(x = Date, y = AveragePrice, color = Type)) + 
  geom_line(aes(x = Date, y = model3_price_prediction)) +
  scale_x_date(date_breaks = "6 month", date_labels =  "%b %Y", expand = c(0,0)) +
  labs(title = "MODEL 3: Multi Linear Regression (TotalVolume + Type + Month + Year + Region)", x = "Date", y = "Average Avocado Price") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, vjust = 0),
        panel.grid.minor = element_line(size = 0.4, linetype = 1),
        panel.grid.major = element_line(size = 0.75, linetype = 3),
        legend.position = "bottom")
```

BLA

```{r fig.height = 10,echo=F, results=T, eval=T, error=F, warning=F, message=F}
# Display multi linear regression plots
grid.arrange(model3_multiLinear.1.plot1, 
             model3_multiLinear.2.plot2, 
             model3_multiLinear.3.plot3, 
             nrow = 3)
```

bla

## 3.4 Model 4: Ridge & Lasso Regression

https://www.statisticshowto.com/lasso-regression/
https://en.wikipedia.org/wiki/Ridge_regression


```{r echo=F, results=T, eval=T, error=F, warning=F, message=F}
# MODEL 4: Ridge regression
# Prepare data
model4.train_data <- train_data %>% select(-c(Date, T4046, T4225, T4770, Day, NonHaasVolume))
model4.test_data <- test_data %>% select(-c(Date, T4046, T4225, T4770, Day, NonHaasVolume))

# Train Model: Ridge Regression (TotalVolume + Type + Month + Year + Region)
model4.1 <- train(AveragePrice ~ ., 
                  data = model4.train_data,
                  method = "ridge")

# Validate model: Ridge Regression (TotalVolume + Type + Month + Year + Region)
model4.1.validation <- model4.1 %>% predict(model4.test_data)
model4.1.RMSE <- RMSE(model4.test_data$AveragePrice, model4.1.validation)
model4.1.R2 <- R2(model4.1.validation, model4.test_data$AveragePrice)

cat("Model 4: R-squared =", model4.1.R2)
cat("Model 4: RMSE =", model4.1.RMSE)

# MODEL 4: Ridge Regression (TotalVolume + Type + Month + Year + Region)
model4.1.plot1 <- test_data %>% 
  mutate(model4_price_prediction = model4.1.validation) %>%
  ggplot() + 
  geom_point(aes(x = Date, y = AveragePrice, color = Type)) + 
  geom_line(aes(x = Date, y = model4_price_prediction)) +
  scale_x_date(date_breaks = "6 month", date_labels =  "%b %Y", expand = c(0,0)) +
  labs(title = "MODEL 4: Ridge Regression (TotalVolume + Type + Month + Year + Region)", x = "Date", y = "Average Avocado Price") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, vjust = 0),
        panel.grid.minor = element_line(size = 0.4, linetype = 1),
        panel.grid.major = element_line(size = 0.75, linetype = 3),
        legend.position = "bottom")

# print model results
model4.1.plot1

# Train Model: Lasso Regression (TotalVolume + Type + Month + Year + Region)
model4.2 <- train(AveragePrice ~ ., 
                  data = model4.train_data,
                  method = "lasso")

# Validate model: Lasso Regression (TotalVolume + Type + Month + Year + Region)
model4.2.validation <- model4.2 %>% predict(model4.test_data)
model4.2.RMSE <- RMSE(model4.test_data$AveragePrice, model4.2.validation)
model4.2.R2 <- R2(model4.2.validation, model4.test_data$AveragePrice)

cat("Model 4: R-squared =", model4.2.R2)
cat("Model 4: RMSE =", model4.2.RMSE)

# MODEL 4: Lasso Regression (TotalVolume + Type + Month + Year + Region)
model4.2.plot1 <- test_data %>% 
  mutate(model4_price_prediction = model4.2.validation) %>%
  ggplot() + 
  geom_point(aes(x = Date, y = AveragePrice, color = Type)) + 
  geom_line(aes(x = Date, y = model4_price_prediction)) +
  scale_x_date(date_breaks = "6 month", date_labels =  "%b %Y", expand = c(0,0)) +
  labs(title = "MODEL 4: Lasso Regression (TotalVolume + Type + Month + Year + Region)", x = "Date", y = "Average Avocado Price") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, vjust = 0),
        panel.grid.minor = element_line(size = 0.4, linetype = 1),
        panel.grid.major = element_line(size = 0.75, linetype = 3),
        legend.position = "bottom")

# print model results
model4.1.plot1

# Train Model: Lasso Regression (TotalVolume + Type + Month + Year + Region)
control <- trainControl(method = "repeatedcv", # k-fold cross-validation
                        number = 10,   # number of folds
                        repeats = 10)  # repeated ten times

# train 
model4.3 <- train(AveragePrice ~ ., 
                  data = model4.train_data,
                  method = "lasso",
                  trControl = control)

# Validate model: Lasso Regression (TotalVolume + Type + Month + Year + Region)
model4.3.validation <- model4.3 %>% predict(model4.test_data)
model4.3.RMSE <- RMSE(model4.test_data$AveragePrice, model4.3.validation)
model4.3.R2 <- R2(model4.3.validation, model4.test_data$AveragePrice)

cat("Model 4: R-squared =", model4.3.R2)
cat("Model 4: RMSE =", model4.3.RMSE)

# MODEL 4: Lasso Regression (TotalVolume + Type + Month + Year + Region)
model4.3.plot1 <- test_data %>% 
  mutate(model4_price_prediction = model4.3.validation) %>%
  ggplot() + 
  geom_point(aes(x = Date, y = AveragePrice, color = Type)) + 
  geom_line(aes(x = Date, y = model4_price_prediction)) +
  scale_x_date(date_breaks = "6 month", date_labels =  "%b %Y", expand = c(0,0)) +
  labs(title = "MODEL 4: Lasso Regression w/ k-fold (TotalVolume + Type + Month + Year + Region)", x = "Date", y = "Average Avocado Price") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, vjust = 0),
        panel.grid.minor = element_line(size = 0.4, linetype = 1),
        panel.grid.major = element_line(size = 0.75, linetype = 3),
        legend.position = "bottom")

# print model results
model4.3.plot1
```

BLA


## 3.5 Model Results Summary

l

# 4. CONCLUSION & DISCUSSION

l

## 4.1 Future considerations

l

## 4.2 References

* Irizarry, Rafael A., “Introduction to Data Science: Data Analysis and Prediction Algorithms in R”, webpage:https://rafalab.github.io/dsbook/
* LaTeX Equation Builder, webpage:https://latex.codecogs.com/eqneditor/editor.php
* knitr::kable and kableExtra Tutorial, webpage:https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
* ggplot2: Quick guide, website:http://www.sthda.com/english/wiki/ggplot2-histogram-plot-quick-start-guide-r-software-and-data-visualization
* Correlograms: website: https://www.statmethods.net/advgraphs/correlograms.html